---
title: "COVID-19 data analysis"
date: '2020-09-15'
description: Deaths by gender and co-morbidities
draft: no
keywords: ''
slug: blog3
categories:
- ''
- ''
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(infer)
```

Let us revisit the [CDC Covid-19 Case Surveillance Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf). There are well over 3 million entries of individual, de-identified patient data. Since this is a large file, I suggest you use `vroom` to load it and you keep `cache=TRUE` in the chunk options.


```{r, cache=TRUE}
# URL link to CDC to download data
url <- "https://data.cdc.gov/api/views/vbim-akqf/rows.csv?accessType=DOWNLOAD"

covid_data <- vroom(url)%>%
  clean_names()


```

Given the data we have, I would like you to produce two graphs that show death % rate:
(Do we need to perform data cleansing on missing values? to be discussed)
(Double faceting for both, use facet_grid, check R docu)
1. by age group, sex, and whether the patient had co-morbidities or not
sex, age_group, medcond_yn, death_yn! 
```{r, fig.width= 20}
covid_data_deathrate1 <- covid_data %>% 
  filter(!(death_yn %in%c("Missing","Unknown")),
         !(medcond_yn %in%c("Missing","Unknown")),
         !(is.na(age_group)), 
         !(age_group %in%c("Missing","Unknown")),
         !(is.na(sex)), 
         !(sex %in%c("Missing","Unknown"))) %>% 
  
 group_by (age_group, sex, medcond_yn, death_yn) %>%
  count(death_yn)  %>%
  group_by (age_group, sex, medcond_yn) %>%
  mutate(death_rate = n*100/sum(n)) %>% 
  filter(death_yn=="Yes") %>% mutate(comorb = ifelse(medcond_yn=="No","Without Comorbidities","With Comorbidities"), death_rate = round(death_rate,1))

covid_data_deathrate1 %>%  
  ggplot(aes(x= age_group, y = death_rate, fill=sex))+
  geom_col()+
  coord_flip()+
  facet_grid(col=vars(sex),row=vars(comorb))+
  theme(axis.text.x = element_text(angle = 30)) +
  geom_text(aes(label=death_rate), position=position_dodge(width=0),hjust = -0.25)+ theme_bw()+
    labs(title="Covid death % by age group, sex and presence of co-morbidities",
       x="",y="", caption="Source:CDC")+
  theme(legend.position = "none")
  


```
1. by age group, sex, and whether the patient was admitted to Intensive Care Unit (ICU) or not.
sex, age_group, icu_yn, death_yn!

```{r covid_challenge_2, echo=FALSE, fig.width=20}
covid_data_deathrate2 <- covid_data %>% 
  filter(!(death_yn %in%c("Missing","Unknown")),
         !(icu_yn %in%c("Missing","Unknown")),
         !(is.na(age_group)), 
         !(age_group %in%c("Missing","Unknown")),
         !(is.na(sex)), 
         !(sex %in%c("Missing","Unknown"))) %>% 
  
 group_by (age_group, sex, icu_yn, death_yn) %>%
  count(death_yn)  %>%
  group_by (age_group, sex, icu_yn) %>%
  mutate(death_rate = n*100/sum(n)) %>% 
  filter(death_yn=="Yes") %>% mutate(ICU = ifelse(icu_yn=="Yes","Admitted to ICU","No ICU"), death_rate = round(death_rate,1))

covid_data_deathrate2 %>%  
  ggplot(aes(x= age_group, y = death_rate, fill=sex))+
  geom_col()+coord_flip()+facet_grid(col=vars(sex),row=vars(ICU))+
  theme(axis.text.x = element_text(angle = 30)) +geom_text(aes(label=death_rate), position=position_dodge(width=0.9),hjust = -0.25)+ ggtitle("Covid death % by age group, sex and whether patient was admitted to ICU")+ theme_bw()+
  labs(title="Covid death % by age group, sex and whether patient was admitted to ICU",
       x="",y="", caption="Source:CDC")+
  theme(legend.position = "none")

